---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  echo = TRUE,
  eval = TRUE
)
```

```{r setup}
library(fishpals)
library(ggplot2)
```

# Introduction

The `fishpals` package provides fish-inspired color palettes and some lightweight `ggplot` themes that we hope will be especially useful to scientists when preparing reports, presentations, and publications.

# 1. Installing and loading `fishpals` into R
<br>

`fishpals` is trivial, so it's unlikely to ever be submitted to CRAN.  You can install it using either `devtools::install_github()` or the `install_github` function from the `remotes` package:

```{r, eval=FALSE}

devtools::install_github("fishsciences/fishpals", build_vignettes = TRUE) 
library(fishpals)

```


# 2. Convenience functions

To see all the palettes included in the `fishpals` package, use `getpals()`:

```{r}
getpals()

```


To view a particular palette, call it by name with `viewpals()`

```{r}
viewpals("surgewrasse")
```

To see a full list of all the colors in each palette, use `get_fishpals_cols()`:

```{r}
get_fishpals_cols()

```


# 3. The `scale_*_fishpals()` functions

`fishpals` is designed to integrate with `ggplot2` code, and so provides two `ggplot2` scale functions for displaying its custom color palettes.  To use them, map a variable to a color aesthetic and call the appropriate scale function: `scale_fill_fishpals()` or `scale_color_fishpals()`:

```{r}

ggplot(warpbreaks) +
   geom_boxplot(aes(x = wool, 
                    y = breaks, 
                    color = wool,
                    fill = wool),
                alpha = 0.5,
                width = 0.5,
                show.legend = FALSE) +
  facet_wrap(~tension,
             labeller = labeller(tension = 
                                   c("L" = "Low",
                                     "M" = "Medium",
                                     "H" = "High")
                              )) +
  scale_color_fishpals(discrete = TRUE) +
  scale_fill_fishpals(discrete = TRUE) +
  labs(x = "Wool type",
       y = "Number of breaks",
       title = "Breaks ~ wool type + tension level")


```

The default color palette is "cfs".  To use a different one, just call it by name:

```{r}
ggplot(warpbreaks) +
   geom_boxplot(aes(x = wool, 
                    y = breaks, 
                    color = wool,
                    fill = wool),
                alpha = 0.5,
                width = 0.5,
                show.legend = FALSE) +
  facet_wrap(~tension,
             labeller = labeller(tension = 
                                   c("L" = "Low",
                                     "M" = "Medium",
                                     "H" = "High")
                              )) +
  scale_color_fishpals("adultchinook", discrete = TRUE) +
  scale_fill_fishpals("adultchinook", discrete = TRUE) +
  labs(x = "Wool type",
       y = "Number of breaks",
       title = "Breaks ~ wool type + tension level")
```


# 4. Lightweight Themes

## `theme_report()`

This theme is designed to have good properties when knitting to pdf or to Microsoft Word.  These properties include:

1. Default size 11 font, with size 11 for axis titles and size 14 for chart titles.
2. Default inner border around the plot margins and the inner panel (`inner_border = TRUE`).
3. Close-ish spacing of facets to fit on a vertical page.
4. Default legend position of bottom, horizontal orientation.


```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  theme_report() +
  labs(y = "Miles per gallon",
       x = "Engine displacement (L)",
       title = "Highway mpg ~ Engine displacement")


ggplot(economics_long, aes(x = date, y = value01)) +
  geom_line(aes(color = variable),
             size = 0.5) +
  theme_report(inner_border = FALSE) +
  scale_color_fishpals() +
  guides(color = guide_legend(title = NULL)) +
  labs(x = "Date",
       y = "Value index",
       title = "Scaled economic indices ~ year") 


ggplot(warpbreaks) +
   geom_boxplot(aes(x = wool, 
                    y = breaks, 
                    color = wool,
                    fill = wool),
                alpha = 0.5,
                width = 0.5,
                show.legend = FALSE) +
  facet_wrap(~tension,
             labeller = labeller(tension = 
                                   c("L" = "Low",
                                     "M" = "Medium",
                                     "H" = "High")
                              )) +
  scale_color_fishpals(discrete = TRUE) +
  scale_fill_fishpals(discrete = TRUE) +
  theme_report() +
  labs(x = "Wool type",
       y = "Number of breaks",
       title = "Breaks ~ wool type and tension level")

```


## `theme_pres()`

This theme is designed with good defaults for presentations.  These include:

1. Default size 14 font
2. Default transparencies for plot.background element_rect()s, which will correspond to a dark theme when you save with ggsave(..., bg = "transparent").
3. Minimalist design for clean slides.
4. Widely-spaced facet panels

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  theme_pres() +
  labs(y = "Miles per gallon",
       x = "Engine displacement (L)",
       title = "Highway mpg ~ Engine displacement")


ggplot(economics_long, aes(x = date, y = value01)) +
  geom_line(aes(color = variable),
             size = 0.5) +
  theme_pres() +
  scale_color_fishpals() +
  guides(color = guide_legend(title = NULL)) +
  labs(x = "Date",
       y = "Value index",
       title = "Scaled economic indices ~ year") 



ggplot(warpbreaks) +
   geom_boxplot(aes(x = wool, 
                    y = breaks, 
                    color = wool,
                    fill = wool),
                alpha = 0.5,
                width = 0.25,
                show.legend = FALSE) +
  facet_wrap(~tension,
             labeller = labeller(tension = 
                                   c("L" = "Low",
                                     "M" = "Medium",
                                     "H" = "High")
                              ),
             nrow = 1) +
  scale_color_fishpals(discrete = TRUE) +
  scale_fill_fishpals(discrete = TRUE) +
  theme_pres() +
  labs(x = "Wool type",
       y = "Number of breaks",
       title = "Breaks ~ wool type + tension level")

```


## `theme_pub()` 

This theme has default font sizes and axis options that are often used in publications.

```{r}

ggplot(economics_long, aes(x = date, y = value01)) +
  geom_line(aes(color = variable),
             size = 0.5) +
  theme_pub() +
  scale_color_brewer(palette = 6) + # using color brewer black and white palette for demonstration
  guides(color = guide_legend(title = NULL)) +
  labs(x = "Date",
       y = "Value index",
       title = "Scaled economic indices ~ year") 

```





