---
title: "fishpals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fishpals}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fishpals)
```

The `fishpals` package provides fish-derived color palettes and some lightweight `ggplot` themes that we hope will be especially useful to scientists when preparing reports, presentations, and publications.

This vignette presents examples of all the palettes and themes included in the package.  It is organized by theme and chart type.

## `theme_report()`

This theme is designed to have good properties when knitting to pdf or to Microsoft Word.  These properties include:

1. Default size 10 font, with size 11 for axis titles and size 13 for chart titles.
2. Border around the plot margins
3. Default legend position of "bottom"
4. Close-ish spacing of facets to fit on a vertical page


```{r}

ggplot(mtcars) +
  geom_point(aes(x = disp, y = hp)) +
  theme_report()


ggplot(iris) +
  geom_bar(aes(x = Species, y = Sepal.Width, fill = Species), stat = "identity") +
  theme_report() +
  scale_fill_fishpals()

ggplot(mtcars) +
   geom_point(aes(x = disp, y = hp, color = factor(carb) ),
              size = 2.5) +
  facet_wrap(~am, scales = "free" ) +
  scale_color_fishpals("surgewrasse", discrete = TRUE) +
  theme_report()


ggplot(mtcars) +
   geom_point(aes(x = disp, y = hp, color = carb ),
              size = 2.5) +
  facet_wrap(~am, scales = "free" ) +
  scale_color_fishpals("surgewrasse", discrete = FALSE) +
  theme_report()


#-------------------------------------------------------#
# install.packages("ggplot2")
# load package and data
options(scipen=999)  # turn-off scientific notation like 1e+48
theme_set(theme_report())  # pre-set the bw theme.
data("midwest", package = "ggplot2")
# midwest <- read.csv("http://goo.gl/G1K41K")  # bkup data source

# Scatterplot
ggplot(midwest, aes(x=area, y=poptotal)) +
  geom_point(aes(col=state, size=popdensity)) +
  geom_smooth(method="loess", se=F) +
  xlim(c(0, 0.1)) +
  ylim(c(0, 500000)) +
  labs(subtitle="Area Vs Population",
       y="Population",
       x="Area",
       title="Scatterplot",
       caption = "Source: midwest") +
  scale_color_fishpals("adultchinook")

#-------------------------------------------------------#

data(mpg, package="ggplot2")

mpg_select <- mpg[mpg$manufacturer %in% c("audi", "ford", "honda", "hyundai"), ]

# Scatterplot

ggplot(mpg_select, aes(displ, cty)) +
  labs(subtitle="mpg: Displacement vs City Mileage",
       title="Bubble chart") +
  geom_jitter(aes(col=manufacturer, size=hwy)) +
  geom_smooth(aes(col=manufacturer), method="lm", se=F) +
  scale_color_fishpals("genidaqs", discrete = TRUE)

#-------------------------------------------------------#
# Data Prep
theme_set(theme_report())
data("mtcars")  # load data
mtcars$`car name` <- rownames(mtcars)  # create new column for car names
mtcars$mpg_z <- round((mtcars$mpg - mean(mtcars$mpg))/sd(mtcars$mpg), 2)  # compute normalized mpg
mtcars$mpg_type <- ifelse(mtcars$mpg_z < 0, "below", "above")  # above / below avg flag
mtcars <- mtcars[order(mtcars$mpg_z), ]  # sort
mtcars$`car name` <- factor(mtcars$`car name`, levels = mtcars$`car name`)  # convert to factor to retain sorted order in plot.

# Diverging Barcharts
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) +
  geom_bar(stat='identity', aes(fill=mpg_type), width=.5)  +
  scale_fill_fishpals("juvsteel",

                    name="Mileage",
                    labels = c("Above Average",
                               "Below Average")) +
  labs(subtitle="Normalised mileage from 'mtcars'",
       title= "Diverging Bars") +
  coord_flip()

```

